\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-03-31 analysis/paper/paper.Rmd
Fixed penal. on time and Brier
-Time is now penalized according to issue #99.
- Brier score is using the riskRegression package now. This uses KM as the censoring distribution... prodlim::prodlim() instead of survival::survfit()

TODO: describe Coefficient plot,

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison, beyond a simple Kaplan-meier curve. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. We will then compare each modeling using the brier score [@graf1999ass] We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps.

The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years.  Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score over all, it will be used in the following comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

```{r baselinePlotsPrep, eval=eval_cs3,echo=FALSE}
completeData <- casebase::support
workingCompleteData <- model.matrix(death ~ . - d.time, data = completeData)[, -c(1)] # remove intercept
# Create u and xCox, which will be used when fitting Cox with glmnet.
newData <- workingCompleteData
x <- workingCompleteData
# x and y will be used to fit the casebase model under glmnet.
# x=as.matrix(sparse.model.matrix(death~ .-d.time+0,data=workingData))
y <- data.matrix(completeData[, c(2, 5)])

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
names(fullDataCox)[names(fullDataCox) == 'V1'] <- 'time'
names(fullDataCox)[names(fullDataCox) == 'V2'] <- 'event'
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

@@

```{r baselines, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:baselines} AbsoluteRisk curves from hazard models. Pink: All covariates without SPS nor APS. Green: APS only model. Blue: SPS only model. Purple: unadjusted curve. All three models underestimate the risk in comparison to the unadjusted curve.",echo=FALSE}
ggplot(as.data.frame(baseCoxComparisonData), aes(Time)) +
  geom_line(aes(y = SPS, colour = "SPS"), lwd = 1) +
  geom_line(aes(y = APS, colour = "APS"), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  geom_line(aes(y = Full, col = "All Covariates without SPS APS"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Days)")
```

```{r brierpt1, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:brier1} Brier scores over time. Full model performs best.",echo=FALSE}

@@

For our penalized logistic regression model, we opted for a logistic function of time, resulting in a Weibull distribution assumption. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where y contains the time and event variables and x contains all other variables we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function.
 We then take our hazard and use it to predict the absolute risk of each individual in our test set, over time. These individuals are then average resulting in the smooth absolute risk curve in Figure \ref{fig:cs3Final}.

**note that I am still penalizing time, there's a bug I did not want to deal with when I try to use `penalty.factor` in fitsmoothhazard . If this is still here, i have not fixed it.**

@@

abCN <- as.data.frame(cbind(abcoxNet$time, 1 - rowMeans(abcoxNet$surv)))
colnames(abCN) <- c("Time", "Absolute risk")
abCN$Model <- "Coxnet"
abCN <- abCN[abCN$Time <= max(abKMcompare[, 1]), ]
# plot(abCN[,1],1-abCN[,2])
```

@@

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 100, standardize = TRUE)
```

@@

```{r include=TRUE,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3Final} HERES SOME TEXT ABT IT"}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 2) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 2) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 2) +
  labs(y = "Probability of Death", x = "Survival time (Days)")
```

???SHORT SPIEL ABOUT IPA BETWEEN SEMI_PARAMETRIC AND PARAMETRIC which leads to conclusion of best overall model. This refers to the COEFPLOT in a following figure???

```{r finalBrierplot, echo=FALSE, eval=eval_cs3}
brierScoreKMCens <- function(predSurvs, times, newData) {
  # inverse probability censoring weights
  # Note that we take probabilities right before the drops.

  #DIFFERENCES START HERE
  fitCens <- prodlim::prodlim(Hist(time, event != 0) ~ 1, newData, reverse = TRUE)
  IPCW.subject.times <- prodlim::predictSurvIndividual(fitCens, lag = 1) # G(t-|X)
  #DIFFERENCES END HERE

  # Empty matrix that will be filled in with the following loop
  Score <- matrix(NA, nrow(predSurvs), ncol(predSurvs))
  IPCWMatrix <- matrix(NA, nrow(predSurvs), ncol(predSurvs))
  # for each point in time we have predicted
  for (i in 1:length(times)) {
    # get number of censored individuals so long as their survival time is less than times[i]
    # these individuals do not have an effect on right-censored brier score.
    CensBefore <- newData$event == 0 & newData$time < times[i]
    # y encompasses all survival times larger than t[i] with a 1, 0 otherwise
    y <- drop(t(newData$time > times[i]))
    # above permits the two parts of right-censored brier score to be calculated, without IPCW, in one line
    Score[i, ] <- (y - predSurvs[i, ])^2

    # Generate IPCWMatrix
    IPCWMatrix[i, y == 0] <- IPCW.subject.times[y == 0] # G(t-|X) filled in corresponding positions
    #DIFFERENCES START HERE
    IPCW.time <- predict(fitCens, newdata = newData, times = times[i], level.chaos = 1, mode = "matrix", type = "surv", lag = 1)
    #DIFFERENCES END HERE
    IPCWMatrix[i, y == 1] <- IPCW.time # G(t) filled, same value, for remaining positions.
    # above calculated individuals who don't have an effect on brier score at specfied time. they're scores are set to 0.
    Score[i, CensBefore] <- 0
  }

  # apply IPCW to all scores
  Err <- Score / IPCWMatrix

  # Average curve demonstrating right-censored brier averaged over test-set, for each time of interest.
  Err <- apply(Err, 1, mean)
  return(Err)
}
absRiskcb <-abCbFull
absRiskcb[, -c(1)] <- 1 - absRiskcb[, -c(1)] # make it survival probabilities
absRiskcb <- absRiskcb[-c(1), -c(1)] # remove extra time at 0. Also remove the first column (times) as it matches perfectly now

kmCensoredBrierScores <- brierScoreKMCens(predSurvs = absRiskcb, times = times, newData = data.frame(time=test$time,event=test$event))

finalBrierDataToPlot<-rbind(
  data.frame(brierScore = brierCoxNet$Brier$score$Brier[brierCoxNet$Brier$score$model == "PredictionModel"],times=times,model="Penalized Cox"),
   data.frame(brierScore = brierFull$Brier$score$Brier[brierFull$Brier$score$model == "PredictionModel"],times=times,model="cox"),
      rbind(data.frame(brierScore=kmCensoredBrierScores,times=times,model="Penalized logistic"))

  )

ggplot(data = finalBrierDataToPlot, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Times") +
  ylab("Brier-Score")+
  ggtitle("Brier scores over time")

>>>>

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison, beyond a Kaplan-meier curve. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years. Therefore, We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. Next, we will compare each modeling using the Brier score [@graf1999ass]. The Brier score is able to demonstrate both discrimination and calibration in a single model [@graf1999ass] and provide a comparable metric between semi-parametric and parametric methods in the survival context. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score on overall, it will be used in the following comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

```{r baselinePlotsPrep, eval=eval_cs3,echo=FALSE}
completeData <- casebase::support

completeData$d.time<-completeData$d.time/365

workingCompleteData <- model.matrix(death ~ . - d.time, data = completeData)[, -c(1)] # remove intercept

# Create u and xCox, which will be used when fitting Cox with glmnet.
newData <- workingCompleteData
x <- workingCompleteData
# x and y will be used to fit the casebase model under glmnet.
# x=as.matrix(sparse.model.matrix(death~ .-d.time+0,data=workingData))
y <- data.matrix(completeData[, c(2, 5)])

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
names(fullDataCox)[names(fullDataCox) == 'V1'] <- 'time'
names(fullDataCox)[names(fullDataCox) == 'V2'] <- 'event'

#here we specify train and test size
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

@@

```{r baselines, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:baselines} AbsoluteRisk curves from hazard models. Pink: All covariates without SPS nor APS. Green: APS only model. Blue: SPS only model. Purple: unadjusted curve. All three models underestimate the risk in comparison to the unadjusted curve.",echo=FALSE}
ggplot(as.data.frame(baseCoxComparisonData), aes(Time)) +
  geom_line(aes(y = SPS, colour = "SPS"), lwd = 1) +
  geom_line(aes(y = APS, colour = "APS"), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  geom_line(aes(y = Full, col = "All Covariates without SPS APS"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)")
```

```{r brierpt1, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:brier1} Brier scores over time. Full model (Cox Full) performs best, while the other two (APS only and SPS only) are comparable.",echo=FALSE}

@@

For our penalized logistic regression model, we opted for a logistic function of time, resulting in a Weibull distribution assumption. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where y contains the time and event variables and x contains all other variables we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function.
 We then take our hazard and use it to predict the absolute risk of each individual in our test set, over time. The probabilites over time for each individual are averaged, resulting in the smooth absolute risk curve in figure \ref{fig:cs3Final}. On average, the Brier score is lower in our penalized logistic regression model, compared to the other models (fig. \ref{fig:cs3FinalBrier}).

This case study serves to demonstrate how regularization can be used in a survival context, using logistic regression. Our approach performs comparably to Cox regression, demonstrating its validity in an applied setting.

@@

abCN <- as.data.frame(cbind(abcoxNet$time, 1 - rowMeans(abcoxNet$surv)))
colnames(abCN) <- c("Time", "Absolute risk")
abCN$Model <- "Penalized Cox"
abCN <- abCN[abCN$Time <= max(abKMcompare[, 1]), ]
# plot(abCN[,1],1-abCN[,2])
```

@@

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 10, standardize = TRUE,penalty.facto=c(0,rep(1,ncol(x))))
```

@@

```{r include=TRUE,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3Final} Four absolute risk curves, integrated from the following hazard models: Cox (pink), penalized Cox (green), penalized logistic (blue) and Kaplan-Meier (purple)."}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 1) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)")
```

```{r riskregressionBrier,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3FinalBrier} 4 curves comparing the brier scores as they change over time between a null model (Kaplan-Meier, black), Cox regression (yellow), penalized Cox regression (blue) and penalized logistic regression (green)."}

predictRisk.singleRisk <- function(object, newdata, times, cause, ...) {
  if (object$call[[1]]!="glm"){
    #get all covariates excluding intercept and time
    coVars=colnames(object$originalData$x)
    #coVars is used in lines 44 and 50
    newdata=data.matrix(drop(subset(newdata, select=coVars)))
  }

  # if (missing(cause)) stop("Argument cause should be the event type for which we predict the absolute risk.")
  # the output of absoluteRisk is an array with dimension dependening on the length of the requested times:
  # case 1: the number of time points is 1
  #         dim(array) =  (length(time), NROW(newdata), number of causes in the data)
  if (length(times) == 1) {
    a <- absoluteRisk(object, newdata = newdata, time = times)
    p <- matrix(a, ncol = 1)
  } else {
    # case 2 a) zero is included in the number of time points
    if (0 %in% times) {
      # dim(array) =  (length(time)+1, NROW(newdata)+1, number of causes in the data)
      a <- absoluteRisk(object, newdata = newdata, time = times)
      p <- t(a)
    } else {
      # case 2 b) zero is not included in the number of time points (but the absoluteRisk function adds it)
      a <- absoluteRisk(object, newdata = newdata, time = times)
      ### we need to invert the plot because, by default, we get cumulative incidence
      #a[, -c(1)] <- 1 - a[, -c(1)]
      ### we remove time 0 for everyone, and remove the time column
      a <- a[-c(1), -c(1)] ### a[-c(1), ] to keep times column, but remove time 0 probabilities
      # now we transpose the matrix because in riskRegression we work with number of
      # observations in rows and time points in columns
      p <- t(a)
    }
  }
  if (NROW(p) != NROW(newdata) || NCOL(p) != length(times)) {
    stop(paste("\nPrediction matrix has wrong dimensions:\nRequested newdata x times: ", NROW(newdata), " x ", length(times), "\nProvided prediction matrix: ", NROW(p), " x ", NCOL(p), "\n\n", sep = ""))
  }
  p
}

class(cbFull) <- c("singleRisk",class(cbFull))
brierFinalResults <- Score(list("Cox" = coxFull,"Penalized Cox"=coxNet,"Penalized Logistic"=cbFull), data = test, formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times =times)

levels(brierFinalResults$Brier$score$model)[levels(brierFinalResults$Brier$score$model)=="Null model"] <- "Kaplan-Meier"
ggplot(data = brierFinalResults$Brier$score, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Times") +
  ylab("Brier-Score")+
  ggtitle("Brier scores over time")

### 2020-03-31 analysis/paper/paper.tex
Fixed penal. on time and Brier
-Time is now penalized according to issue #99.
- Brier score is using the riskRegression package now. This uses KM as the censoring distribution... prodlim::prodlim() instead of survival::survfit()

TODO: describe Coefficient plot,

Before generating an absolute risk curve using regularized logistic
regression, we will explore three options as a baseline comparison,
beyond a simple Kaplan-meier curve. For all the following analyses in
this case study, we will use 80\% of the observations to train our
models, while 20\% will be used to generate absolute risk curves, all of
which will be averaged to demonstrate each model's predictive ability.
We will then compare each modeling using the brier score
\citep{graf1999ass} We will examine three cox regression scenarios: sps
only, aps only and all covariates excluding sps and aps.

The original study compared these scores over a span of 180 days, which
may not be comparable over the full 5.56 years. Over 5.56 years,
\ref{fig:baselines} demonstrates that all the cox models have similar,
if not identical average predictions. Figure \ref{fig:brier1} examines
the brier scores as they change over time. As the full model has the
lowest brier score over all, it will be used in the following comparison
to regularized cox, regularized logistic regression and the unadjusted
Kaplan-Meier curve.

@@

\caption{\label{fig:brier1} Brier scores over time. Full model performs best.}\label{fig:brierpt1}
\end{figure}
\end{CodeChunk}

@@

For our penalized logistic regression model, we opted for a logistic
function of time, resulting in a Weibull distribution assumption. As
casebase makes use of the glmnet package, we interact with the
fitsmoothhazard.fit function using a matrix interface, where y contains
the time and event variables and x contains all other variables we would
like to include in the model. For the SUPPORT dataset, all factor
variables were converted into dummy variables. Here, we used lasso
penalization by setting alpha to 1, leading to our smooth hazard
function. We then take our hazard and use it to predict the absolute
risk of each individual in our test set, over time. These individuals
are then average resulting in the smooth absolute risk curve in Figure
\ref{fig:cs3Final}.

\textbf{note that I am still penalizing time, there's a bug I did not
want to deal with when I try to use \texttt{penalty.factor} in
fitsmoothhazard . If this is still here, i have not fixed it.}

@@

\begin{CodeInput}
R> 
R> cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 100, standardize = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

\caption{\label{fig:cs3Final} HERES SOME TEXT ABT IT}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

???SHORT SPIEL ABOUT IPA BETWEEN SEMI\_PARAMETRIC AND PARAMETRIC which
leads to conclusion of best overall model. This refers to the COEFPLOT
in a following figure???

\begin{center}\includegraphics{../figures/finalBrierplot-1} \end{center}

@@

This report was generated on 2020-03-26 16:26:24 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.1 (2019-07-05)
#> Platform: x86_64-w64-mingw32/x64 (64-bit)
#> Running under: Windows 10 x64 (build 18363)
#> 
#> Matrix products: default
#> 
#> locale:
#> [1] LC_COLLATE=English_United States.1252 
#> [2] LC_CTYPE=English_United States.1252   
#> [3] LC_MONETARY=English_United States.1252
#> [4] LC_NUMERIC=C                          
#> [5] LC_TIME=English_United States.1252    
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.5               dotwhisker_0.5.0         
#>  [3] prodlim_2019.11.13        riskRegression_2020.02.05
#>  [5] pracma_2.2.9              kableExtra_1.1.0         
#>  [7] data.table_1.12.8         glmnet_3.0-2             
#>  [9] Matrix_1.2-17             casebase_0.2.1.9001      
#> [11] survival_3.1-8            magrittr_1.5             
#> [13] forcats_0.5.0             stringr_1.4.0            
#> [15] dplyr_0.8.3               purrr_0.3.3              
#> [17] readr_1.3.1               tidyr_1.0.2              
#> [19] tibble_2.1.3              ggplot2_3.3.0            
#> [21] tidyverse_1.3.0          
#> 
#> loaded via a namespace (and not attached):
#>  [1] nlme_3.1-140        fs_1.3.2            cmprsk_2.2-9       
#>  [4] lubridate_1.7.4     rms_5.1-4           webshot_0.5.2      
#>  [7] RColorBrewer_1.1-2  httr_1.4.1          numDeriv_2016.8-1.1
#> [10] tools_3.6.1         backports_1.1.5     R6_2.4.1           
#> [13] rpart_4.1-15        Hmisc_4.3-1         DBI_1.1.0          
#> [16] mgcv_1.8-28         colorspace_1.4-1    nnet_7.3-12        
#> [19] withr_2.1.2         gridExtra_2.3       tidyselect_1.0.0   
#> [22] compiler_3.6.1      quantreg_5.54       cli_2.0.2          
#> [25] rvest_0.3.5         htmlTable_1.13.3    SparseM_1.78       
#> [28] xml2_1.2.5          sandwich_2.5-1      labeling_0.3       
#> [31] checkmate_2.0.0     scales_1.1.0        mvtnorm_1.1-0      
#> [34] polspline_1.1.17    digest_0.6.25       foreign_0.8-71     
#> [37] rmarkdown_2.1       base64enc_0.1-3     rticles_0.14       
#> [40] jpeg_0.1-8.1        pkgconfig_2.0.3     htmltools_0.4.0    
#> [43] dbplyr_1.4.2        htmlwidgets_1.5.1   rlang_0.4.5        
#> [46] readxl_1.3.1        rstudioapi_0.11     VGAM_1.1-2         
#> [49] farver_2.0.3        shape_1.4.4         generics_0.0.2     
#> [52] zoo_1.8-7           jsonlite_1.6.1      acepack_1.4.1      
#> [55] Formula_1.2-3       Rcpp_1.0.3          munsell_0.5.0      
#> [58] fansi_0.4.1         lifecycle_0.2.0     multcomp_1.4-12    
#> [61] stringi_1.4.6       yaml_2.2.1          MASS_7.3-51.4      
#> [64] plyr_1.8.6          ggstance_0.3.3      grid_3.6.1         
#> [67] crayon_1.3.4        lattice_0.20-38     haven_2.2.0        
#> [70] hms_0.5.3           timereg_1.9.4       knitr_1.28         
#> [73] pillar_1.4.3        codetools_0.2-16    stats4_3.6.1       
#> [76] reprex_0.3.0        glue_1.3.1          evaluate_0.14      
#> [79] latticeExtra_0.6-29 modelr_0.1.6        vctrs_0.2.4        
#> [82] png_0.1-7           foreach_1.4.8       MatrixModels_0.4-1 
#> [85] cellranger_1.1.0    gtable_0.3.0        assertthat_0.2.1   
#> [88] xfun_0.12           viridisLite_0.3.0   iterators_1.0.12   
#> [91] cluster_2.1.0       lava_1.6.7          TH.data_1.0-10
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master C:/Users/jesse/Documents/gits/cbpaper
#> Remote:   master @ origin (https://github.com/Jesse-Islam/cbpaper.git)
#> Head:     [a01a2a8] 2020-03-06: Fixed lollipop plot
\end{CodeOutput}
\end{CodeChunk}

>>>>

Before generating an absolute risk curve using regularized logistic
regression, we will explore three options as a baseline comparison,
beyond a Kaplan-meier curve. The original study compared these scores
over a span of 180 days, which may not be comparable over the full 5.56
years. Therefore, We will examine three cox regression scenarios: sps
only, aps only and all covariates excluding sps and aps. Next, we will
compare each modeling using the Brier score \citep{graf1999ass}. The
Brier score is able to demonstrate both discrimination and calibration
in a single model \citep{graf1999ass} and provide a comparable metric
between semi-parametric and parametric methods in the survival context.
For all the following analyses in this case study, we will use 80\% of
the observations to train our models, while 20\% will be used to
generate absolute risk curves, all of which will be averaged to
demonstrate each model's predictive ability. Over 5.56 years,
\ref{fig:baselines} demonstrates that all the cox models have similar,
if not identical average predictions. Figure \ref{fig:brier1} examines
the brier scores as they change over time. As the full model has the
lowest brier score on overall, it will be used in the following
comparison to regularized cox, regularized logistic regression and the
unadjusted Kaplan-Meier curve.

@@

\caption{\label{fig:brier1} Brier scores over time. Full model (Cox Full) performs best, while the other two (APS only and SPS only) are comparable.}\label{fig:brierpt1}
\end{figure}
\end{CodeChunk}

@@

For our penalized logistic regression model, we opted for a logistic
function of time, resulting in a Weibull distribution assumption. As
casebase makes use of the glmnet package, we interact with the
fitsmoothhazard.fit function using a matrix interface, where y contains
the time and event variables and x contains all other variables we would
like to include in the model. For the SUPPORT dataset, all factor
variables were converted into dummy variables. Here, we used lasso
penalization by setting alpha to 1, leading to our smooth hazard
function. We then take our hazard and use it to predict the absolute
risk of each individual in our test set, over time. The probabilites
over time for each individual are averaged, resulting in the smooth
absolute risk curve in figure \ref{fig:cs3Final}. On average, the Brier
score is lower in our penalized logistic regression model, compared to
the other models (fig.~\ref{fig:cs3FinalBrier}).

This case study serves to demonstrate how regularization can be used in
a survival context, using logistic regression. Our approach performs
comparably to Cox regression, demonstrating its validity in an applied
setting.

@@

\begin{CodeInput}
R> 
R> cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 10, standardize = TRUE,penalty.facto=c(0,rep(1,ncol(x))))
\end{CodeInput}
\end{CodeChunk}

@@

\caption{\label{fig:cs3Final} Four absolute risk curves, integrated from the following hazard models: Cox (pink), penalized Cox (green), penalized logistic (blue) and Kaplan-Meier (purple).}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

{\centering \includegraphics{../figures/riskregressionBrier-1}

\caption{\label{fig:cs3FinalBrier} 4 curves comparing the brier scores as they change over time between a null model (Kaplan-Meier, black), Cox regression (yellow), penalized Cox regression (blue) and penalized logistic regression (green).}\label{fig:riskregressionBrier}
\end{figure}
\end{CodeChunk}

@@

This report was generated on 2020-03-31 18:19:20 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.1 (2019-07-05)
#> Platform: x86_64-w64-mingw32/x64 (64-bit)
#> Running under: Windows 10 x64 (build 18363)
#> 
#> Matrix products: default
#> 
#> locale:
#> [1] LC_COLLATE=English_United States.1252 
#> [2] LC_CTYPE=English_United States.1252   
#> [3] LC_MONETARY=English_United States.1252
#> [4] LC_NUMERIC=C                          
#> [5] LC_TIME=English_United States.1252    
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.5               dotwhisker_0.5.0         
#>  [3] prodlim_2019.11.13        riskRegression_2020.02.05
#>  [5] pracma_2.2.9              kableExtra_1.1.0         
#>  [7] data.table_1.12.8         glmnet_3.0-2             
#>  [9] Matrix_1.2-17             casebase_0.2.1.9001      
#> [11] survival_3.1-8            magrittr_1.5             
#> [13] forcats_0.5.0             stringr_1.4.0            
#> [15] dplyr_0.8.3               purrr_0.3.3              
#> [17] readr_1.3.1               tidyr_1.0.2              
#> [19] tibble_2.1.3              ggplot2_3.3.0            
#> [21] tidyverse_1.3.0          
#> 
#> loaded via a namespace (and not attached):
#>  [1] nlme_3.1-140        fs_1.3.2            cmprsk_2.2-9       
#>  [4] lubridate_1.7.4     rms_5.1-4           webshot_0.5.2      
#>  [7] RColorBrewer_1.1-2  httr_1.4.1          numDeriv_2016.8-1.1
#> [10] tools_3.6.1         backports_1.1.5     R6_2.4.1           
#> [13] rpart_4.1-15        Hmisc_4.3-1         DBI_1.1.0          
#> [16] mgcv_1.8-28         colorspace_1.4-1    nnet_7.3-13        
#> [19] withr_2.1.2         gridExtra_2.3       tidyselect_1.0.0   
#> [22] compiler_3.6.1      quantreg_5.54       cli_2.0.2          
#> [25] rvest_0.3.5         htmlTable_1.13.3    SparseM_1.78       
#> [28] xml2_1.2.5          sandwich_2.5-1      labeling_0.3       
#> [31] checkmate_2.0.0     scales_1.1.0        mvtnorm_1.1-0      
#> [34] polspline_1.1.17    digest_0.6.25       foreign_0.8-71     
#> [37] rmarkdown_2.1       base64enc_0.1-3     rticles_0.14       
#> [40] jpeg_0.1-8.1        pkgconfig_2.0.3     htmltools_0.4.0    
#> [43] dbplyr_1.4.2        htmlwidgets_1.5.1   rlang_0.4.5        
#> [46] readxl_1.3.1        rstudioapi_0.11     VGAM_1.1-2         
#> [49] farver_2.0.3        shape_1.4.4         generics_0.0.2     
#> [52] zoo_1.8-7           jsonlite_1.6.1      acepack_1.4.1      
#> [55] Formula_1.2-3       Rcpp_1.0.4          munsell_0.5.0      
#> [58] fansi_0.4.1         lifecycle_0.2.0     multcomp_1.4-12    
#> [61] stringi_1.4.6       yaml_2.2.1          MASS_7.3-51.4      
#> [64] plyr_1.8.6          ggstance_0.3.3      grid_3.6.1         
#> [67] crayon_1.3.4        lattice_0.20-38     haven_2.2.0        
#> [70] hms_0.5.3           timereg_1.9.4       knitr_1.28         
#> [73] pillar_1.4.3        codetools_0.2-16    stats4_3.6.1       
#> [76] reprex_0.3.0        glue_1.3.1          evaluate_0.14      
#> [79] latticeExtra_0.6-29 modelr_0.1.6        vctrs_0.2.4        
#> [82] png_0.1-7           foreach_1.4.8       MatrixModels_0.4-1 
#> [85] cellranger_1.1.0    gtable_0.3.0        assertthat_0.2.1   
#> [88] xfun_0.12           viridisLite_0.3.0   iterators_1.0.12   
#> [91] cluster_2.1.0       lava_1.6.7          TH.data_1.0-10
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master C:/Users/jesse/Documents/gits/cbpaper
#> Remote:   master @ origin (https://github.com/Jesse-Islam/cbpaper.git)
#> Head:     [5f22217] 2020-03-26: AddedBrier
\end{CodeOutput}
\end{CodeChunk}
